Smoking and COVID Living Review (v12): a bayesian analysis

### 2021-07-23 reports/bayesian_report.Rmd
begin v12

---
title: 'Smoking and COVID Living Review (v10): a bayesian analysis'
author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:

@@

```{r head_search_details}
flextable(search_details[,1:13])
```
The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.

This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`. These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
```{r tail_study_details, include=FALSE}
datatable(data_study_general[1:4] %>%
            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
          escape = F)
```
The third sheet produces **Table 1** this contains the details of the population of the studies in our review. Primarily includes smoking status but also the proportion of females in the study and the age of participants.
```{r tail_table_1, include=FALSE}
datatable(table_1 %>%
                 filter(lead_author != is.na(lead_author)))
```
The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status. Most studies will not have data on this outcome however we filter these subsequently. These are sparse datasets as many studies do not report on all levels of smoking exposure.
```{r tail_table_2, include=FALSE}
datatable(
  table_2 %>%
  filter(data_on_testing == TRUE)
)
```
The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
```{r tail_table_3, include=FALSE}
datatable(
  table_3 %>%
  filter(data_on_hospitalisation == TRUE)
)
```
The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status. Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
```{r tail_table_4, include=FALSE}
datatable(
  table_4 %>%
  filter(data_disease_severity == TRUE)
)
```
The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status. A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
```{r tail_table_5, include=FALSE}
datatable(
  table_5 %>%
  filter(data_on_deaths == TRUE)
)
```
The final sheet is used to collate data on **Study quality**. The method of assessing quality changed in version 3 of the review. We now use the following criteria:

@@

``` {r excluding_studies}
superseded_studies <- names(tidyselect::vars_select(data_study_general$lead_author, ends_with("_old", ignore.case = T)))
old_isaric <- head(names(tidyselect::vars_select(data_study_general$lead_author, starts_with("isaric", ignore.case = T))), n = -1)
#This will prevent duplication of the ISARIC reports, only retaining the last based on their names
excluded_other_reason <- c("mehra", #retracted
                           "hopkinson", #self-reported
                           "gaibazzi", #withdrawn
                           "kua" #self-reported
                           )

@@

This is a searchable table.
``` {r descriptives}
source(here("scripts", "remove_excluded.r"))
current_total <- nrow(a)
datatable(a)

@@

## The PRISMA diagram will be updated
The PRISMA reporting guidelines for systematic reviews the guidance is available [here](http://www.prisma-statement.org/). This takes the format of the number of articles obtained from the search from [OVID](https://ovidsp.dc1.ovid.com/sp-4.07.0b/ovidweb.cgi) (login may be needed), [medRxiv](https://www.medrxiv.org/search/text_abstract_title%3Asmoking%2BAND%2BCOVID%20text_abstract_title_flags%3Amatch-all%20jcode%3Amedrxiv%20numresults%3A75%20sort%3Apublication-date%20direction%3Adescending%20format_result%3Astandard), and articles identified by the research team. These are then screened at the title and abstract level. Full texts are then read to extract relevant data, if articles are excluded at this level we describe the reason for this. All of the selected articles will contribute to the narrative synthesis but only those that are judged 'good' or 'fair' quality will be included in the Bayesian meta-analysis.
``` {r PRISMA}
previous_total <- 345

@@

``` {r countries, cache = TRUE}
#Countries
country <- table_1$country %>%
  to_upper_camel_case() %>%
  recode(., 'Usa' = 'USA', 'Uk' = 'UK', 'SaudiArabia' = 'Saudi Arabia', "SouthKorea" = "South Korea", "UnitedArabEmirates" = "United Arab Emirates") %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Country' = 1,
         'Number' = 2) %>%
  mutate(Country = as.character(Country))

@@

country <- subset(country, Country != "Multiple")
```
Currently studies are reported from `r paste(nrow(country))`
``` {r, cache=TRUE}
country <- mutate_geocode(country, Country)

country_bounds <- country %>%
  select(lon, lat) %>%
  tidyr::drop_na() %>%
  sp::SpatialPointsDataFrame(coords = .[,1:2], data = .) %>%
  sp::bbox() * 1.1

>>>>

---
title: 'Smoking and COVID Living Review (v12): a bayesian analysis'
author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:

@@

```{r head_search_details}
flextable(search_details %>%
            mutate(`date_screening in format (2020-05-22)` = ymd(`date_screening in format (2020-05-22)`)))
```
The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.

This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`. These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
```{r tail_study_details, include=FALSE}
datatable(data_study_general[1:4] %>%
            mutate(`date_published_format(2020-05-12)` = ymd(data_study_general$`date_published_format(2020-05-12)`),
                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
          escape = F)
```
The third sheet produces **Table 1** this contains the details of the population of the studies in our review. Primarily includes smoking status but also the proportion of females in the study and the age of participants.
```{r tail_table_1, include=FALSE}
datatable(table_1 %>%
                 filter(lead_author != is.na(lead_author)))
```
The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status. Most studies will not have data on this outcome however we filter these subsequently. These are sparse datasets as many studies do not report on all levels of smoking exposure.
```{r tail_table_2, include=FALSE}
datatable(
  table_2 %>%
  filter(data_on_testing == TRUE)
)
```
The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
```{r tail_table_3, include=FALSE}
datatable(
  table_3 %>%
  filter(data_on_hospitalisation == TRUE)
)
```
The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status. Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
```{r tail_table_4, include=FALSE}
datatable(
  table_4 %>%
  filter(data_disease_severity == TRUE)
)
```
The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status. A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
```{r tail_table_5, include=FALSE}
datatable(
  table_5 %>%
  filter(data_on_deaths == TRUE)
)
```
The final sheet is used to collate data on **Study quality**. The method of assessing quality changed in version 3 of the review. We now use the following criteria:

@@

``` {r excluding_studies}
superseded_studies <- data_study_general %>%
  filter(str_ends(lead_author, "_old")) %>%
  pull(lead_author)

old_isaric <- data_study_general %>%
  filter(str_starts(lead_author, "isaric")) %>%
  head(., -1) %>%
  pull(lead_author)
#This will prevent duplication of the ISARIC reports, only retaining the last based on their names

excluded_other_reason <- c("mehra", #retracted
                           "hopkinson", #self-reported
                           "gaibazzi", #withdrawn
                           "kua" #self-reported
                           )

@@

This is a searchable table.
``` {r descriptives}
source(here("scripts", "remove_excluded.r")) # This removes the excluded studies from each table that has been read in previously.
current_total <- nrow(a)
datatable(a)

@@

## The PRISMA diagram will be updated
The PRISMA reporting guidelines for systematic reviews the guidance is available [here](http://www.prisma-statement.org/). This takes the format of the number of articles obtained from the search from [OVID](https://ovidsp.dc1.ovid.com/sp-4.07.0b/ovidweb.cgi) (login may be needed), [medRxiv](https://www.medrxiv.org/search/text_abstract_title%3Asmoking%2BAND%2BCOVID%20text_abstract_title_flags%3Amatch-all%20jcode%3Amedrxiv%20numresults%3A75%20sort%3Apublication-date%20direction%3Adescending%20format_result%3Astandard), and articles identified by the research team. These are then screened at the title and abstract level. Full texts are then read to extract relevant data, if articles are excluded at this level we describe the reason for this. All of the selected articles will contribute to the narrative synthesis but only those that are judged 'good' or 'fair' quality will be included in the Bayesian meta-analysis.
``` {r PRISMA}
previous_total <- read_rds(here("data_clean", "data_study_general.rds")) %>%
  filter(review_version %in% prev_versions) %>%
  nrow()

@@

``` {r countries, cache = TRUE}
#Countries
country <- table_1$country %>%
  to_upper_camel_case(sep_out = " ") %>%
  recode(., 'Usa' = 'USA', 'Uk' = 'UK') %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Country' = 1,
         'Number' = 2)

@@

country <- country %>%
  filter(Country != "Multiple")
```
Currently studies are reported from `r paste(nrow(country))` countries.

``` {r, cache=TRUE}
countries <- ne_countries(type = "countries", returnclass = "sf")

map <- country %>%
  mutate(iso3c = countrycode::countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
  left_join(., countries %>%
              as_tibble() %>%
              dplyr::select(adm0_a3, geometry), by = c("iso3c" = "adm0_a3")) %>%
  st_as_sf()

country_bounds <- map %>%
  st_bbox()
